Handling String-Number Conversion in String Constraint Solving

### 2019-11-09 draft.tex
I am done, and Yu-Fang promissed that he will not bother me from this point on.

\newcommand{\nat}{\mathbb{N}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\pim}{\alpha}
\newcommand{\syncop}{\Cap}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#(#1)}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}

@@

We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the sets of natural numbers (including 0) and 
integers. For a set $A$, we use $|A|$ to denote its size. 
An \emph{alphabet} is a finite set $\Sigma$ of \emph{characters} and a \emph{string} over $\Sigma$ is a sequence $w = a_1\ldots a_n$ of characters from $\Sigma$, with $\epsilon$ denoting the \emph{empty string}. 
We use $w_1\cdot w_2$ to denote the \emph{concatenation} of strings $w_1$ and $w_2$.
$\Sigma^*$ is the set of all strings over $\Sigma$, $\Sigma^+ = \Sigma^*\setminus \{\epsilon\}$, and 
A \emph{language} over $\Sigma$ is a subset $L$ of $\Sigma^*$.  
%
We use $|w|$ to denote the length of $w$ and $|w|_a$ to denote the number of occurrences of the character $a\in \Sigma$ in $w$.

@@

@@

In this section, we formally define string constraints. To begin with, we fix an finite alphabet $\Sigma \subseteq \mathbb{N}$. Note that here we assume the alphabet is a finite subset of natural numbers. Essentially, we try to capture the numerical encoding of the corresponding symbols in computers (e.g., in ASCII, `A' is encoded as $65$). We assume there is an one-one mapping between numbers in $\Sigma$ and the character it encodes. For the simplicity of presentation, we assume the character `$0$' is mapped to the number $0$, `$1$' to $1$,$\ldots$, and `$9$' to $9$. For other character $c$, we use $\enc{c}$ to denote the number that it maps to. Notice that this approach is general enough to support any finite set of characters. We write $[n,m]$ for the set of numbers (or symbols in $\Sigma$) $\{k\mid n\leq k \leq m\}$.

@@

We introduce \emph{parametric flat automata} (PFA) that will be used to define patterns 
%for the string variable domain restriction. 
used by the under-approximation module to restrict the domain of string variables. 
Namely, an PFA is a finite automaton $A = (Q,T,\cvar,\st{0}{0},\st{m}{0})$ operating over an alphabet $\cvar$ of \emph{character variables}, 
and satisfying the following structural constraints:
\begin{enumerate}
	\item The final state $\st{m}{0}$ is reached from  the initial state $\st{0}{0}$ through a stright path of $m-1$ transitions $(\st{i}{0},v_i,\st{i+1}{0}) \in T$, $\st{i}{0} \in Q$ for $i\in[0,m-1]$. %, with pairwise distiand states $\st{i}{0} \in Q$ for $i\in[0,m-1]$.
	\item  
Each state $\st{i}{0}$ is the origin of a unique simple cycle of the length $l_i\in\mathbb{N}$, consisting of states $\st{i}{j-1} \in Q$ and transitions $(\st{i}{j-1}, \sym{i}{j-1}, \st{i}{j \bmod l_i})$ for $j\in [1,l_i]$. 
Notice that case when $l_i = 0$ is also admissible and means that there is no cycle on $q_i$.
%Assume that $\st{i}{0}$ is included in a cycle with $l_i$ states (includes $\st{i}{0}$). 
%For each $j\in [1,l_i]$, we have $\st{i}{j} \in Q$ and $(\st{i}{j-1}, \sym{i}{j-1}, \st{i}{j \bmod l_i})$. 
%(where $l_i = 0$ means that there is no cycle on $q_i$)
	\item Each variable in $\cvar$ appears on at most one transition.% (but $\epsilon$ transition can appear multiple times). 
\end{enumerate} 
\lh{TODO: decide whether we want epsilon trans, or perhaps additional constraints saying that a var equals epsilon or a character}
\begin{figure*}
	\tikzset{state/.style={circle,draw=blue!50,fill=blue!20,
			thick,inner sep=0pt,minimum size=6mm}, initial text=$ $}
	\begin{tikzpicture} 
		\node[state,initial] (q0)  {$\st{0}{0}$};
		\node[state] (q1) [right = 3cm of q0] {$\st{1}{0}$};
		\node (q2) [right = 1.8cm of q1]{};
		\node (q3) [right = 1cm of q2]{};
		\node[state,accepting] (qm) [right = 1.8cm of q3] {$\st{m}{0}$};

@@

\begin{lemma}\label{lemma:parikh}
For an PFA $A$ and $x,y\in L(A)$ if $\parikhof x = \parikhof y$ then $x = y$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$. \lh{say precisely how?}
%
For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
We may say that $\parikhof x$ is an encoding of $x$, and define the decoding function $\decode A$ such that $\decode A(\parikhof x) = x$.

@@

Since every string $w \in \semof{A}$ is obtained from a parametric string $x\in L(A)$ and an interpretation of character varriables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric string $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different strings are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup x) = w$.
Encodings of strings in $\semof{A}$ are encodings of parametric strings of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$, 
hence $\semof A$ is characterised by models of $\parikhof {A}$ interpreted over variables $\cvar\cup\pvar$ (variables of $\cvar$  may have arbitrary values since $\parikhof {A}$ does not constraint them), we may say that 
%\begin{lemma}\ref{lemma:decw}
%For a PFA $A$, $\semof{A} = \{\decode(I\cup \pim) \mid \pim\in\semof{\parikhfof A} \land I:\cvar\rightarrow \Sigma_\epsilon\}$.
%\end{lemma}
$$
\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}
$$
where $\semof{\parikhfof A}_{\cvar\cup\pvar}$ are models of $\parikhfof A$, assignments to $\pvar$, united with arbitrary interpretations of $\cvar$ \lh{this is ugly}\yfc{I totally agree}.

\paragraph{Synchronisation of PFA}
The following construction of the so called synchronisation formula of two PFA $A$ and $A'$ is a crucial component of building underapproximations of string constraints. 
Synchronisation formula is a formula $\syncfof A {A'}$ over $\pvarone\cup\pvartwo\cup\cvarone\cup \cvartwo$ encoding the semantic intersection $\semof A \cap \semof {A'}$. Besides the intersection, it contains a more precise information about the synchronisation of $A$ and ${A'}$, 
namely, it characterises the set of pairs of encodings $(I\cup \pim)\in \parikhof {A}$ and $(I'\cup \pim')\in\parikhof {A'}$ such that 
$\decode A(I\cup \pim) = \decode {A'}(I'\cup \pim')$.% which constitute the intersection.

Assume that $A = (Q,T,\cvarone,q_i,q_f)$ and ${A'} = (Q',T',\cvartwo,q_i',q_f')$ are two SFA over disjoint sets of variables, $\cvarone \cap \cvartwo = \emptyset$. 
We build so called asynchronous product automaton $\syncof A {A'}$ of $A$ and ${A'}$ and then extract the formula $\syncfof A {A'}$ from its Parikh image.
%
Such product automaton uses $Q\times Q'$ as the set of states and 
%$(\cvarone \cup \{\epsilon\}) \times (\cvartwo \cup \{\epsilon\} )$ 
$\cvarone_\epsilon\times\cvartwo_\epsilon$
as the alphabet. 
%
Every accepting run of $\syncof A {A'}$ corresponds to an accepting run of $A$ and an accepting run of ${A'}$, and the word accepted by the run of $\syncof A {A'}$ implies a set of constraints on $\pvarone\cup\pvartwo\cup\cvarone\cup\cvartwo$ under which the word accepted by $A$ has the same semantics as the word accepted by ${A'}$.
%
Particularly, if a transition $((q_1,q_1'), (v,v'),(q_2,q_2'))$ is taken, it means the character variable $v$ and $v'$ should be assigned the same value and this value leads $A$ from state $q_1$ to state $q_2$ and ${A'}$ from $q_1'$ to $q_2'$.
%
If a transition $((q_1,q_1'), (v,\epsilon),(q_2,q_1'))$ is taken, then it means that the character variable $v$ should be assigned $\epsilon$, $A$ may transition under $v$ to $q_1$ and ${A'}$ stays in the same state, as a response to $A$s reading epsilon may as well be taking no action. Symmetrically, ${A'}$ might read a variable $v$ assigned $\epsilon$ and $A$ may stay.

Formally, the product automaton is a tuple $\syncof A {A'} = (Q\times Q', \syncT, \cvarone_\epsilon \times \cvartwo_\epsilon, (q_i,q_i'),(q_f,q_f'))$, where the transition relation $\syncT$ is the minimal set satisfying the following:
%
\begin{itemize}
\item If $(q_1,v,q_2) \in T$ and $(q_1',v',q_2') \in T'$, then we have $((q_1,q_1'),(v,v'),(q_2,q_2'))\in \syncT$.
\item If $(q_1,v,q_2) \in T$, then for all states $q'\in Q'$, we have $((q_1,q'),(v,\epsilon),(q_2,q'))\in \syncT$.
\item If $(q_1',v',q_2') \in T'$, then for all states $q\in Q$, we have $((q,q_1'),(\epsilon,v'),(q,q_2'))\in \syncT$.
\end{itemize}	
%
Next, we define the set of constraints that extract from $\syncof A {A'}$ pairs of united assignments (encodings of strings over $\Sigma$) that encode the same string. 
We start with computing the Parikh formula $\parikhfof {\syncof A {A'}}$ of $\syncof A {A'}$ over the set of variables $\pvarsof{\cvarone_\epsilon\times\cvartwo_\epsilon}$ and then conjoin it with the following restricting constraints. 
%We use $\#v$ and $\#v'$ to denote the number of occurrences of $v_i$ and $v_i^{(j)}$, respectively.

The first restricting constraint establishes the relation of between the Parikh image of a word accepted by $\syncof A {A'}$ and the Parikh images of the corresponding parametric strings accepted by $A$ and ${A'}$:
$$ \Psi_{\#} = 
\left(\bigwedge_{v\in\cvarone}\#v = \sum_{x' \in \cvartwo_\epsilon} \#{(v,x')}\right)
\land
\left(\bigwedge_{v'\in\cvartwo} \#v' = \sum_{x \in \cvarone_\epsilon} \#{(x,v')}\right)
$$
Notice that $x,x'$ are either variables or $\epsilon$.
%\begin{enumerate}
%	\item for all $v \in \cvarone$, $\#(v) = \sum_{v' \in \cvartwo} \#{(v,v')}$   
%	\item for all $v \in \cvartwo$, $\#(v) = \sum_{v' \in \cvarone} \#{(v',v)}$ 
%\end{enumerate}

Then we add the following linear constraints to enforce that the decodings of the models are indeed the same, 
% condition 1. and 2. required for the pair of symbolic string $(w_n, w_m)$ and $I$, i.e., $I(w_n) = I(w_m)$ and $I$ assign the same value to all versions of the same variable (we use $-1$ as the value of $\epsilon$). 
$$ \Psi_= =
\bigwedge_{x\in\cvarone_\epsilon,x'\in\cvartwo_\epsilon} \#(x,x')>0 \rightarrow (x=x')
$$
%\begin{enumerate}
%	\item for all $v\in \cvarone$ and $v'\in \cvartwo$, $\#{(v,v')}>0 \rightarrow (v=v')$, 
%	\item for all $v\in \cvarone$, $\#{(v,\epsilon)}>0 \rightarrow (v=\epsilon)$, and
%	\item for all $v\in \cvartwo$,$\#{(\epsilon,v')}>0 \rightarrow (v'=\epsilon)$.
%\end{enumerate}

The synchronisation formula is the conjunction 
$$\label{lemma:synccorrect}
%\syncfof A {A'} = \exists \pvarsof{\cvarone\times\cvartwo}\exists {(\cvarone\times\cvartwo)}.\,\parikhfof{\syncof A {A'}} \land \Psi_{\#} \land \Psi_=
\syncfof A {A'} = {(\cvarone\times\cvartwo)}.\,\parikhfof{\syncof A {A'}} \land \Psi_{\#} \land \Psi_=
$$
The correctness of this construction is expressed by Lemma~\ref{lemma:synccorrect}. Informally,
the

\begin{lemma}
 $\semof {\syncfof A {A'}} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
\end{lemma}

>>>>

\newcommand{\nat}{\mathbb{N}}
\newcommand{\todo}[1]{{\color{blue}TODO: #1}}
\newcommand{\lh}[1]{{\color{orange}Lukas: #1}}
\newcommand{\yfc}[1]{{\color{blue}YFC: #1}}
\newcommand{\petr}[1]{{\color{pink}Petr: #1}}
\newcommand{\sti}[1]{\mbox{\textsf{toInt}($#1$)}}
\newcommand{\its}[1]{\mbox{\textsf{toStr}($#1$)}}
\newcommand{\varn}{\mbox{$\mathbb{V}_{\mathbb{Z}}$}}
\newcommand{\vars}{\mbox{$\mathbb{V}_{\Sigma^*}$}}
\newcommand{\cvars}{\mbox{$\mathbb{V}_{\Sigma_\epsilon}$}}
\newcommand{\pvars}{\mbox{$\mathbb{V}_{\sharp}$}}
%\newcommand{\cvar}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvarone}{\mbox{$V_{\Sigma_\epsilon}$}}
%\newcommand{\cvartwo}{\mbox{$V'_{\Sigma_\epsilon}$}}
\newcommand{\cvarone}{V}
\newcommand{\cvartwo}{V'}
\newcommand{\cvar}{V}
\newcommand{\modelsof}[1]{[\![#1]\!]}
\newcommand{\true}{\mbox{$\mathsf{true}$}}
\newcommand{\false}{\mbox{$\mathsf{false}$}}
\newcommand{\enc}[1]{[\![#1]\!]}
%\newcommand{\parikhof}[1]{\mathit{Parikh}{(#1)}}
\newcommand{\parikhof}[1]{\mathbb{P}{(#1)}}
\newcommand{\parikhwof}[2]{|#1|_{#2}}
%\newcommand{\semof}[1]{||#1||}
\newcommand{\semof}[1]{\modelsof{#1}}
\newcommand{\decode}[1]{\mathit{decode_{#1}}}
\newcommand{\parikhfof}[1]{\Phi_{\parikhof{#1}}}
\newcommand{\pim}{\alpha}
%\newcommand{\syncop}{\Cap}
\newcommand{\syncop}{\times}
\newcommand{\syncof}[2]{#1 \syncop #2}
\newcommand{\syncfof}[2]{\Psi_{\syncof {#1} {#2}}}
\newcommand{\syncT}{T_\syncop}
\newcommand{\pvarsof}[1]{\#{#1}}
\newcommand{\pvar}{\pvarsof V}
\newcommand{\pvarone}{\pvarsof V}
\newcommand{\pvartwo}{\pvarsof {V'}}
\newcommand{\defeq}{::=}

@@

We use $\mathbb{N}$ and $\mathbb{Z}$ to denote the sets of natural numbers (including 0) and 
integers. For a set $A$, we use $|A|$ to denote its size. 
For $n,m\in\nat$, we write $[n,m]$ for the set of natural numbers 
%(or symbols in $\Sigma$) 
$\{k\mid n\leq k \leq m\}$. 
An \emph{alphabet} is a finite set $\Sigma$ of \emph{characters} and a \emph{string} over $\Sigma$ is a sequence $w = a_1\ldots a_n$ of characters from $\Sigma$, with $\epsilon$ denoting the \emph{empty string}. 
We use $w_1\cdot w_2$ to denote the \emph{concatenation} of strings $w_1$ and $w_2$.
$\Sigma^*$ is the set of all strings over $\Sigma$, $\Sigma^+ = \Sigma^*\setminus \{\epsilon\}$, $\Sigma_\epsilon = \Sigma\cup\{\epsilon\}$, and 
A \emph{language} over $\Sigma$ is a subset $L$ of $\Sigma^*$.  
%
We use $|w|$ to denote the length of $w$ and $|w|_a$ to denote the number of occurrences of the character $a\in \Sigma$ in $w$.

@@

Through the paper, we will use quantifier-free linear integer arithmetic formulae, and call them \emph{linear formulae} for short.   
Given a linear formula $\phi$, a set of variables $V$ that includes all variables occurring in $\phi$, and an \emph{interpretation} of $V$, a function $I:V\rightarrow \nat$, 
we denote by $I\models_V \phi$ that $I$ satisfies $\phi$ (which is defined in the standard manner), and call $I$ a model of $\phi$ over $V$.
We use $\modelsof \phi$ to denote the set all models of $\phi$ over $V$. 
We drop the reference to $V$ and write just ``model of $\phi$'',  $\modelsof \phi$, and $\models$ when no confusion may arise.

@@

In this section, we formally define string constraints. To begin with, we fix an finite alphabet $\Sigma \subseteq \mathbb{N}$. Note that here we assume the alphabet is a finite subset of natural numbers. Essentially, we try to capture the numerical encoding of the corresponding symbols in computers (e.g., in ASCII, `A' is encoded as $65$). We assume there is an one-one mapping between numbers in $\Sigma$ and the character it encodes. For the simplicity of presentation, we assume the character `$0$' is mapped to the number $0$, `$1$' to $1$,$\ldots$, and `$9$' to $9$. For other character $c$, we use $\enc{c}$ to denote the number that it maps to. Notice that this approach is general enough to support any finite set of characters.

@@

We introduce \emph{parametric flat automata} (PFA) that will be used to define patterns 
%for the string variable domain restriction. 
used by the under-approximation module to restrict the domain of string variables. 
Namely, an PFA is a finite automaton $A = (Q,T,\cvar,\st{0}{0},\st{m}{0})$ operating over an alphabet $\cvar$ of \emph{character variables}, 
and satisfying the following structural constraints:
\begin{enumerate}
	\item The final state $\st{m}{0}$ is reached from  the initial state $\st{0}{0}$ through a stright path of $m-1$ transitions $(\st{i}{0},v_i,\st{i+1}{0}) \in T$, $\st{i}{0} \in Q$ for $i\in[0,m-1]$. %, with pairwise distiand states $\st{i}{0} \in Q$ for $i\in[0,m-1]$.
	\item  
Each state $\st{i}{0}$ is the origin of a unique simple cycle of the length $l_i\in\mathbb{N}$, consisting of states $\st{i}{j-1} \in Q$ and transitions $(\st{i}{j-1}, \sym{i}{j-1}, \st{i}{j \bmod l_i})$ for $j\in [1,l_i]$. 
Notice that case when $l_i = 0$ is also admissible and means that there is no cycle on $q_i$.
%Assume that $\st{i}{0}$ is included in a cycle with $l_i$ states (includes $\st{i}{0}$). 
%For each $j\in [1,l_i]$, we have $\st{i}{j} \in Q$ and $(\st{i}{j-1}, \sym{i}{j-1}, \st{i}{j \bmod l_i})$. 
%(where $l_i = 0$ means that there is no cycle on $q_i$)
	\item Each variable in $\cvar$ appears on at most one transition.% (but $\epsilon$ transition can appear multiple times). 
\end{enumerate} 
\begin{figure*}
	\tikzset{state/.style={circle,draw=blue!50,fill=blue!20,
			thick,inner sep=0pt,minimum size=6mm}, initial text=$ $}
	\begin{tikzpicture} 
		\node[state,initial] (q0)  {$\st{0}{0}$};
		\node[state] (q1) [right = 3cm of q0] {$\st{1}{0}$};
		\node (q2) [right = 1.8cm of q1]{};
		\node (q3) [right = 1cm of q2]{};
		\node[state,accepting] (qm) [right = 1.8cm of q3] {$\st{m}{0}$};

@@

\begin{lemma}\label{lemma:parikh}
For an PFA $A$ and $x,y\in L(A)$, if $\parikhof x = \parikhof y$ then $x = y$. 
\end{lemma}
%
Indeed, since every variable appears on at most one transition, then the $\parikhof x$ value of all variables appearing within the same cycle is the same, and it is equal to the number of repetitions of that cycle in the accepting run. 
%
The accepting run on $x$ (and so $x$ itself) can thus be reconstructed from $\parikhof x$. \lh{say precisely how?}
%
For example, in the automaton of Figure~\ref{fig:sfa_def}, 
from $\parikhwof x {v_0} = \parikhwof x {v_1} = \cdots = \parikhwof x {v_{m-1}}=1$ and $\parikhwof x {v^0_1}=\parikhwof x {v^1_1}=\parikhwof x {v^2_1}=2$ we derive that $x=v_0(v^0_1v^1_1v^2_1)^2v_1\cdots v_{m-1}$. 
We may say that $\parikhof x$ is an encoding of $x$, and define the decoding function $\decode A$ such that $\decode A(\parikhof x) = x$.

@@

Since every string $w \in \semof{A}$ is obtained from a parametric string $x\in L(A)$ and an interpretation of character varriables $I:\cvar\rightarrow\Sigma_\epsilon$ such that $w = I(x)$, and since $x = \decode A(\parikhof x)$, 
we may say that the assignment $I \cup \parikhof x$ uniting the interpretation $I$ and the Parikh image of the parametric string $x$ encodes $w$. Lemma~\ref{lemma:parikh} also implies that two different strings are encoded by different united assignments (the Parikh image or the interpretation must be different), and hence we may extend the decoding function $\decode A$ to united assignments so that $\decode A(I \cup x) = w$.
Encodings of strings in $\semof{A}$ are encodings of parametric strings of $L(A)$ paired with arbitrary interpretations of character variables $\cvar$, which are precisely all modesl of the formula $\parikhfof A$ interpreted over $\cvar\cup\pvarsof \cvar$, as formulated byt the following collolary of Lemma~\ref{lemma:decx}:
%
\begin{corollary}
For an PFA $A$ over character variables $V$,
$\semof{A} = \{\decode A(I\cup \pim) \mid I\cup\pim\in\semof{\parikhfof A}_{\cvar\cup\pvar}\}$.
\end{corollary}
%
A minor technical difficulty is that we wish to work with pure arithmetic formulae, 
but we may also wish to specify when a character variable is assigned $\epsilon$, which is not a number. 
Hence, to allow purely arithmetic formulae to talk about the value $\epsilon$, we encode $\epsilon$ as $-1$.
In the further development, $\epsilon$ will mean $-1$ when working with arithmetic formulae constraining interpretations of character variables.

\paragraph{Synchronisation of PFA}
The following construction of the so called synchronisation formula of two PFA $A$ and $A'$ over character variables $V$ and $V'$, respectively, is a crucial component of building underapproximations of string constraints. 
Synchronisation formula is a formula $\syncfof A {A'}$ over $\cvarone\cup \cvartwo \cup \pvarone\cup\pvartwo$ encoding the semantic intersection $\semof A \cap \semof {A'}$. Its representation of the intersection includes an additional information about the synchronisation of $A$ and ${A'}$, also needed for building underapproximations of string constraints. 
Namely, it characterises the set of pairs of string encoding encodings $(I\cup \pim)\in \parikhof {A}$ and $(I'\cup \pim')\in\parikhof {A'}$ such that 
$\decode A(I\cup \pim) = \decode {A'}(I'\cup \pim')$.% which constitute the intersection.

Let $A = (Q,T,\cvarone,q_i,q_f)$ and ${A'} = (Q',T',\cvartwo,q_i',q_f')$ such that $\cvarone \cap \cvartwo = \emptyset$. 
We will build their asynchronous product, an automaton $\syncof A {A'}$, 
and then extract the synchronisation formula $\syncfof A {A'}$ from it.
%
The product automaton uses $Q\times Q'$ as the set of states and 
$\cvarone_\epsilon\times\cvartwo_\epsilon$
as the alphabet. 
%
Every accepting run of $\syncof A {A'}$ will correspond to a pair of accepting runs, 
a run of $A$ over a parametric string $x$ and a run of ${A'}$ over a parametric string $x'$. 
The word accepted by the run of $\syncof A {A'}$ will induce the constraints on the interpretations of $I$ of $\cvarone$ and $I'$ of $\cvartwo$ under which the two parametric strings have the same interpretation, i.e. $I(x) = I(x')$.

Intuitively, when $\syncof A {A'}$ takes a transition $((q_1,q_1'), (v,v'),(q_2,q_2'))$, 
it means the character variable $v$ and $v'$ should be assigned the same value,
$A$ transitions under $v$ from state $q_1$ to state $q_2$ and ${A'}$ from $q_1'$ to $q_2'$ under $v'$.
%
When $\syncof A {A'}$ takes a transition $((q_1,q_1'), (v,\epsilon),(q_2,q_1'))$, 
it means that the character variable $v$ should be assigned $\epsilon$, 
$A$ transitions under $v$ to $q_1$, and ${A'}$ takes no action, since no action is needed to match $A$s reading $\epsilon$ (hence not reading the input word). Symmetrically, ${A'}$ might read a variable $v$ assigned $\epsilon$ and $A$ may stay.

Formally, the product automaton is a tuple $\syncof A {A'} = (Q\times Q', \syncT, \cvarone_\epsilon \times \cvartwo_\epsilon, (q_i,q_i'),(q_f,q_f'))$, where the transition relation $\syncT$ is the minimal set satisfying the following:
%
\begin{itemize}
\item If $(q_1,v,q_2) \in T$ and $(q_1',v',q_2') \in T'$, then we have $((q_1,q_1'),(v,v'),(q_2,q_2'))\in \syncT$.
\item If $(q_1,v,q_2) \in T$, then for all states $q'\in Q'$, we have $((q_1,q'),(v,\epsilon),(q_2,q'))\in \syncT$.
\item If $(q_1',v',q_2') \in T'$, then for all states $q\in Q$, we have $((q,q_1'),(\epsilon,v'),(q,q_2'))\in \syncT$.
\end{itemize}	
%
%The we define of constraints that extract from $\syncof A {A'}$ pairs of united assignments (encodings of strings over $\Sigma$) that encode the same string. 
The core part of the synchronisation formula is the Parikh formula $\parikhfof {\syncof A {A'}}$ of the product, which encodes all its runs.
%
To extract form a run of $\syncof A {A'}$ the corresponding run of $A$ and of $A'$, we use the following constraint:
$$ \Psi_{\#} \defeq 
\left(\bigwedge_{v\in\cvarone}\#v = \sum_{x' \in \cvartwo_\epsilon} \#{(v,x')}\right)
\land
\left(\bigwedge_{v'\in\cvartwo} \#v' = \sum_{x \in \cvarone_\epsilon} \#{(x,v')}\right)
$$
Notice that $x,x'$ are either variables or $\epsilon$.
%\begin{enumerate}
%	\item for all $v \in \cvarone$, $\#(v) = \sum_{v' \in \cvartwo} \#{(v,v')}$   
%	\item for all $v \in \cvartwo$, $\#(v) = \sum_{v' \in \cvarone} \#{(v',v)}$ 
%\end{enumerate}
Finally, to ensure that the interpretations of the parametric strings accepted by $A$ and $A'$ are the same, the following constraint is used:
% condition 1. and 2. required for the pair of symbolic string $(w_n, w_m)$ and $I$, i.e., $I(w_n) = I(w_m)$ and $I$ assign the same value to all versions of the same variable (we use $-1$ as the value of $\epsilon$). 
$$ \Psi_= \defeq
\bigwedge_{x\in\cvarone_\epsilon,x'\in\cvartwo_\epsilon} \#(x,x')>0 \rightarrow (x=x')
$$
%\begin{enumerate}
%	\item for all $v\in \cvarone$ and $v'\in \cvartwo$, $\#{(v,v')}>0 \rightarrow (v=v')$, 
%	\item for all $v\in \cvarone$, $\#{(v,\epsilon)}>0 \rightarrow (v=\epsilon)$, and
%	\item for all $v\in \cvartwo$,$\#{(\epsilon,v')}>0 \rightarrow (v'=\epsilon)$.
%\end{enumerate}
The synchronisation formula is the conjunction 
$$\label{lemma:synccorrect}
%\syncfof A {A'} = \exists \pvarsof{\cvarone\times\cvartwo}\exists {(\cvarone\times\cvartwo)}.\,\parikhfof{\syncof A {A'}} \land \Psi_{\#} \land \Psi_=
\syncfof A {A'} \defeq \parikhfof{\syncof A {A'}} \land \Psi_{\#} \land \Psi_=
$$
The correctness of this construction is expressed by Lemma~\ref{lemma:synccorrect}. 
%
%\newcommand\restrict[1]{\raisebox{-.5ex}{$\big|$}_{#1}}
\begin{lemma}
%$\semof {\exists V_\times . \syncfof A {A'}}\restrict{\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
$\semof {\exists V_\times . \syncfof A {A'}} = \{(I\cup\pim)\cup(I'\cup\pim') \mid \decode A(I\cup\pim) = \decode {A'}(I'\cup\pim')\}$. 
\end{lemma}
%
Informally, the models of $\syncfof A {A'}$ encode precisely the pairs of equivalent encodings of words from $\semof{A}$ and $\semof{B}$, that constitute the intersection $\semof{A} \cap \semof{B}$. A minor technical difficulty is that the modes of $\syncfof A {A'}$ include also an assignment to the auxiliary variables of $V_\times = (\cvarone_\epsilon\times\cvartwo_\epsilon) \cup \pvarsof{(\cvarone_\epsilon\times\cvartwo_\epsilon)}$, hence in the lemma we have to restrict the models only to the original variables $\cvarone \cup \cvartwo \cup \pvarsof{\cvarone} \cup \pvarsof\cvartwo$, which we achieve by quantifying out $V_\times$.
